### 2007-01-19 Documentation/Manual.tex
*** empty log message ***

\vspace*{-4ex}
\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible. The software
as well as up-to-date versions of this manual can be found at this URL:
 \verb+http://www.mpi-hd.mpg.de/~globes+

\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependence, such as supernov\ae, can not be studied. It 
is, however, possible to simulate beams with bunch structure, since the 
time dependence of the neutrino source is physically only important to suppress backgrounds.

@@

@@

When using the built-in Earth matter density profile, the 
original source is \Ref~\cite{Stacey}.

All files ending with \verb+.dat+ or \verb+.glb+ 
in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which works should be
cited when using a certain file. Make sure that dependencies are correctly
tracked, \ie , in some cases files included by other files need to be 
checked, too (for example, cross section or flux files). One can use 
the \verb+-v3+ option to \verb+globes+ to see which files
are included.

It is recommended that you use the same style for your own input files, since then, in case they are distributed, everybody will know how to correctly reference your work.

@@

\begin{itemize}
 \item
  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Sec~\ref{sec:userchi}
and~\ref{sec:rules}
 \item
  User-defined priors to include arbitrary external information in the $\chi^2$ before
marginalization over the oscillation parameters; see \Sec~\ref{sec:userdefined} 
\item
  Non-standard physics support; see \Chapt~\ref{chapt:nsphysics}
\item
  Beta beam fluxes available as built-in fluxes; see \Sec~\ref{sec:source}
\item
  Enhanced support for parallelization, such as Condor; see, \eg, page~\pageref{sec:condor}
\item
  Updated AEDL files; see \Tab~\ref{tab:experiments}
\item
 New \AEDL\ features, such as the support of lists as variables and an interpolation
routine; see \Sec~\ref{sec:advaedl}
\item
 Clean-up of inconsistencies, such as an overall (internal) normalization factor in the flux files;
see, \eg, page~\pageref{app:flux}
\item
  Faster probability engine, easier installation (internal changes)
\item
  Experimental feature: Alternative minimizer provided, which is usually faster than
the standard minimizer; see \Chapt~\ref{chapt:experimental}
\end{itemize}

@@

Most of the modifications should not require that old pieces of software be changed. 
However, the following changes are relevant:
\begin{itemize}
\item
 AEDL requires now that {\tt \$version} be used to define the minimum version number this AEDL
file is to be used with. With this requirement one can easily avoid that new AEDL
files with new features be used with old versions of \GLOBES\ which may not recognize these features.
\item
 Some of the earlier AEDL files have been updated, changed names, or have been removed. In addition, new files have been added. Although old AEDL files will run as usual for compatibility, they will not be
supported by the \GLOBES\ team anymore. You should make sure to keep these files when updating \GLOBES . 
\end{itemize}

@@

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES .
After that, the user
interface is successively introduced from very basic to more sophisticated
functions. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.

@@

@@

This work and the development of \GLOBES\ have over the years been supported by:
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\end{itemize}

@@

\printindex{aedl}{AEDL reference}

>>>>

\vspace*{-4ex}
\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino oscillation 
long baseline and reactor experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language (\AEDL), which allows to describe most classes 
of long baseline experiments
at an abstract level. On the other hand, it provides a C-library to
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for GNU/Linux. Since the source code is included,
the port to other operating systems is in principle possible. The software
as well as up-to-date versions of this manual can be found at this URL:
 \verb+http://www.mpi-hd.mpg.de/~globes+

\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependence, such as supernov\ae, can not be studied. It 
is, however, possible to simulate beams with bunch structure, since the 
time dependence of the neutrino source is physically only important to suppress backgrounds. 
Furthermore, experiments with discrete numbers of sources and detectors can
be implemented by user-defined systematics in \GLOBES\ 3.0 and higher.

@@

In the newest version \GLOBES\ 3.0 flexibility is introduced at all levels. At the probability level,
the transition probabilities can be modified to introduce new physics. At the systematics level,
user-defined systematical errors and correlations between sources or detectors can be simulated,
and at the analysis level, arbitrary  input from external measurements can be added.
Therefore, \GLOBES\ now provides solutions for new classes of problems.

@@

When using the built-in Earth matter density profile, the 
original source is \Refs~\cite{Dziewonski:1981xy,Stacey}.

All files ending with \verb+.dat+ or \verb+.glb+ 
in the \verb+data+ subdirectory of the \GLOBES\ tar-ball have on top a comment field which clearly indicates which studies should be
cited when using a certain file. Make sure that dependencies are correctly
tracked, \ie , in some cases files included by other files need to be 
checked, too (for example, cross section or flux files). One can use 
the \verb+-v3+ option to the \verb+globes+ command to see which files
are included (\cf, \Chapt~\ref{chap:exp_def}).
It is recommended that you use the same style for your own input files, since then, in case they are distributed, everybody will know how to correctly reference your work.

@@

\begin{itemize}
 \item
  User-defined systematics, which can be used to simulate reactor experiments \etc ; see \Secs~\ref{sec:userchi}
and~\ref{sec:rules}
 \item
  User-defined priors to include arbitrary external information in the $\chi^2$ before
marginalization over the oscillation parameters; see \Sec~\ref{sec:userdefined} 
\item
  Non-standard physics support; see \Chapt~\ref{chapt:nsphysics}
\item
  Beta beam fluxes available as built-in fluxes; see \Sec~\ref{sec:source}
\item
  Enhanced support for parallelization, such as Condor; see, \eg, page~\pageref{sec:condor}
\item
  Updated \AEDL\ files; see \Tab~\ref{tab:experiments}
\item
 New \AEDL\ features, such as the support of lists as variables and an interpolation
routine; see \Sec~\ref{sec:advaedl}
\item
 Clean-up of inconsistencies, such as an overall (internal) normalization factor in the flux files;
see, \eg, page~\pageref{app:flux}
\item
  Faster probability engine, easier installation (internal changes)
\item
  Experimental feature: Alternative minimizer provided, which is usually faster than
the standard minimizer; see \Chapt~\ref{chapt:experimental}
\end{itemize}

@@

Most of the modifications should not require that old pieces of software be changed. 
However, the following changes could be relevant:
\begin{itemize}
\item
 \GLB{SetDensityParams} has to be used with \GLB{DefineParams} and \GLB{SetDensityProjectionFlag} together
with \GLB{DefineProjection}, because unexpected pre-defined behavior should be avoided.
\item
 Functions \GLB{SetFilter}, \GLB{GetFilter}, \GLB{SetFilterState}, and \GLB{GetFilterState} replaced
by functions {\tt ...InExperiment}.
\item
 \AEDL\ requires now that {\tt \$version} be used to define the minimum version number this \AEDL\
file is to be used with. With this requirement one can easily avoid that new \AEDL\
files with new features be used with old versions of \GLOBES\ which may not recognize these features.
\item
 Some of the earlier \AEDL\ files have been updated, changed names, or have been removed. In addition, new files have been added. Although old \AEDL\ files will run as usual for compatibility, they will not be
supported by the \GLOBES\ Team anymore. You should make sure to keep these files when updating \GLOBES . 
\item 
 The implementation of the tilt (systematics) has slightly changed. The tilt also works for
variable bin widths. Therefore, you will obtain slightly different results when you run
the same \AEDL\ between older and newer versions of \GLOBES .
\end{itemize}

@@

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES .
After that, the user
interface is successively introduced from very basic to more sophisticated
functions. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can obtain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.
The corresponding figures are produced with the
Mathematica Notebook {\tt DocPlots.nb}, which can be found
in the example directory as well.

@@

In this version of the manual, introductory topics and advanced topics are
mixed if they belong to the same subject. Therefore, we have marked more
advanced material by a star~($^*$). This material can be skipped in a first
reading of the manual. In some cases, it may be even recommendable to do
so because knowledge of \AEDL\ is required (which is introduced in
the second part).

@@

This work and the development of \GLOBES\ have over the years been supported by
(in chronological order):
\begin{itemize}
\item Technische Universit\"at M\"unchen [All authors]
\item Max-Planck-Institut f\"ur Physik, M\"unchen [PH]
\item Sonderforschungsbereich 375 f\"ur Astro-Teilchenphysik\\
der Deutschen Forschungsgemeinschaft [All authors]
\item Studienstiftung des Deutschen Volkes [JK, WW]
\item Institute for Advanced Study, Princeton [WW]
\item W. M. Keck Foundation [WW]
\item National Science Foundation [WW]
\item University of Wisconsin, Madison [PH]
\item Max-Planck-Institut f\"ur Kernphysik, Heidelberg [JK, ML]
\item Emmy Noether-Programm der Deutschen Forschungsgemeinschaft [WW]
\item Universit\"at W\"urzburg [WW]
\end{itemize}

@@

\printindex{aedl}{\AEDL\ reference}
